---
output: github_document
---

```{r}
if(!require(librarian)){install.packages("librarian")}
librarian::shelf(tidyverse, janitor)
fs::file_delete(dir(here::here("output"), full.names = T))
```



# Replacement Workflow/Basic Data Cleaning



```{r}
ipa_01 <- 
  c(
    "hhid" # label
    , "read"
    , "write" # numeric
    , 'head_gender' #0  female
    , "head_education"
    , "size"
  )
ipa_01_out <- here::here("output", "hh_round1_clean.dta")
```


```{r}
ipa_d_01 <- 
  here::here('input', "hh_round1.dta") %>% haven::read_dta() %>% 
  relocate(ipa_01)
names(ipa_d_01)
```


```{r}
ipa_d_01 %>% 
  distinct(head_education) %>% 
  mutate(
    lab = head_education %>% haven::as_factor("labels")
    , num = head_education %>% haven::as_factor("values")
    , lab1 = paste(lab, "=", num)
  ) %>% 
  arrange(num) %>% 
  pull(lab1)

head_education_labels <- c(
 none = 0                
 ,"nursery/ pre-unit" = 1   
 ,'std 1' = 2               
 ,'std 2' = 3               
 ,'std 3' = 4               
 ,'std 4' = 5               
 ,'std 5' = 6               
 ,'std 6' = 7               
 ,'std 7' = 8               
 ,'std 8' = 9               
 ,'form 1' = 10             
 ,'form 2' = 11             
 ,'form 3' = 12             
 ,'form 4' = 13             
 ,'form 5' = 14             
 ,'form 6' = 15             
 ,'uni. 1' = 16             
 ,'uni. 2' = 17            
 ,'uni. 3' = 18            
 ,'uni. 4' = 19             
 ,'uni. 5' = 20             
 ,"vocational training" = 2
 ,"adult education" = 22    
 ,"not indicated" = 98      
 ,other = 23              
 ,rta = 99                
 ,"NA" = NA
)

```


```{r}


ipa_out <- ipa_d_01 %>% 
  mutate(
    across(.cols = c(read, write, head_gender, head_education), as.numeric)
    , hhid1 = str_sub(hhid, 2, -1) %>% as.numeric()
    , head_gender = ifelse(hhid1 == 15064, 0, head_gender)
    , head_gender = (ifelse(head_gender == 2, 0, 1)) %>% haven::labelled(c(Female = 0, Male = 1)) 
    , head_education = (ifelse(hhid1 == 29038, 0, head_education))
    , read = ifelse(hhid1 == 53024, 3, read)
    ) %>% 
  rename(hh_size = size) %>% 
  select(!hhid1)
    
ipa_out  
```


2.	What percent of respondents have a female head of household?
7.	The Principal Investigator wants to see if there is a relationship between household head gender and household size. Find the average household size by household head gender. 

```{r}
percent <- 
  ipa_out %>% 
  count(head_gender) %>% 
  mutate(
    head_gender = haven::as_factor(head_gender, "labels")
    , n = (((n / sum(n)) * 100) %>% round(2) ) %>% paste("%")
  ) %>% 
  rename(
    "%" = 2
  )
percent
```


Respuesta: `r percent[1, 2][[1]]`

```{r}
ipa_out %>% mutate(
  
    head_gender = haven::as_factor(head_gender, "labels")
) %>% 
  select(hh_size, head_gender) %>% 
  group_by(head_gender) %>% 
  summarise(hh_size_mean = mean(hh_size)) 
```


```{r}
ipa_out %>% haven::write_dta(ipa_01_out)
```


```{r}
rm(list = ls())
```

# Import and Export Excel



```{r}
excel_out <- here::here("output", "excel_dup.xlsx")
```


```{r}
ipa_02 <- 
  readxl::read_xlsx(here::here("input", "round2.xlsx"), sheet = 2, skip = 2) %>% 
  mutate(hhid = as.character(hhid))
```

```{r}
dup <- 
  ipa_02 %>% 
  count(hhid) %>% 
  filter(n > 1) 
```

```{r}
ipa_02_1 <- 
  ipa_02 %>% 
  filter(hhid %in% pull(dup, hhid))
```

```{r}
ipa_02_2 <- 
  ipa_02 %>% 
  pull(hhid)
```



```{r}
xlsx::write.xlsx(
  ipa_02_1
  , file = excel_out
  , sheetName = "duplicados"
  , append =  T
)
xlsx::write.xlsx(
  ipa_02_2
  , file = excel_out
  , sheetName = "duplicados_list"
  , append =  T
)
```

```{r}
rm(list = ls())
```


# Survey Design Reconciliation


```{r}
join1 <- here::here("input", "hh_nr_round1.dta") %>% haven::read_dta()
join2 <- here::here("input", "hh_round1.dta") %>% haven::read_dta()
```

## Formato largo

```{r}
join1_1 <- 
  join1 %>% 
  separate_rows(crop_l, convert = T) %>% 
  rename(cropl = crop_l)
join1_1
```



```{r}
join2_1 <- 
  join2 %>% 
  pivot_longer(contains("cropl")) %>% 
  filter(value > 0) %>% 
  relocate(name) %>% 
  select(!value) %>% 
  rename(cropl = name) %>% 
  mutate(
    cropl = parse_number(cropl)
  )
join2_1
```


```{r}
full_join(join2_1, join1_1) %>% 
  mutate(
    cropl = haven::labelled(
      cropl
      , c(
        	"Rice" = 1
        	, "Cassava" = 2
        	, "Millet" = 3
        	, "Groundnut" = 4
        	, "Sweet Potato" = 5
        	, "Wheat" = 6
        	, "Sorghum" = 7
      )
    )
  ) %>% 
  mutate(
    across(.cols = c(read, write, head_gender, head_education), as.numeric)
    , hhid1 = str_sub(hhid, 2, -1) %>% as.numeric()
    , head_gender = ifelse(hhid1 == 15064, 0, head_gender)
    , head_gender = (ifelse(head_gender == 2, 0, 1)) %>% haven::labelled(c(Female = 0, Male = 1)) 
    , head_education = (ifelse(hhid1 == 29038, 0, head_education))
    , read = ifelse(hhid1 == 53024, 3, read)
    ) %>% 
  rename(hh_size = size) %>% 
  select(!hhid1) %>% 
  haven::write_dta(here::here("output", "hh_round1_longer.dta"))
```
